---
title: "BaseLine"
author: "Marcos Mariscal"
date: "23/5/2021"
output: html_document
---
```{r setup, include=FALSE}

# Getting project directory 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
proydir <- rprojroot::find_rstudio_root_file()

```

#### This RMarkdown gaves first aproach with a lineal regression model

```{r call_libraries, include=FALSE}

# Calling R libraries
library(tidyverse)
#library(psych)
library(caret)
library(Metrics)

```


```{r envdefs, include=FALSE}
#Environment directories vars
airfilepath="./Data/Air_Quality/"
pngpath="./png/"
datapath="./Data/"

# Defining file names variables to load the data
file_TotRespData = "TotRespData.Rdata"

```



```{r Load_TotAgg, include=FALSE}
# Read AQI Visits
load(file = paste(datapath,file_TotRespData, sep=""))

```

Getting only numeric Features for Respiratory illnesses

```{r numeric}

NumDataResp <- TotRespData[,8:26]

```

---

### Normalization the Data

---

#### Z-score normalization, consists of subtracting the mean and divide by the standard deviation
 We normalize the data to bring all the variables to the same range

```{r norm}
means <- colMeans(NumDataResp)
sds <- apply(NumDataResp, 2, sd)

z_NumDataResp <- scale(NumDataResp, center = means, scale = sds)

z_NumDataResp <- as_tibble(z_NumDataResp)


```

#### Splitting data into test and train
The following code splits 70% of the data selected randomly into training set and the remaining 30% sample into test data set

```{r splitting}

dt = sort(sample(nrow(NumDataResp), nrow(NumDataResp)*.7))
train<-z_NumDataResp[dt,]
test<-z_NumDataResp[-dt,]
```

### BASE LINE.Linear Regression
The fist approach for a good prediction model y to get results of a basic model such a Linear Regression

```{r linregression1}

linreg1<- lm(AQI~., data = train)
summary(linreg1)

```
#### We have to coefficients smaller than 0.05 in Features Ages_18_44 and RPS010, then theses are good features to predict Bad Air Quality conditions.

Regarding R squared value is not good 0.20, because it refers that this model has good predictions only for the 20% of the values.
Some variables could be taken out to for this model.


```{r linregression2}

linreg2<- lm(AQI~Ages_0_4 +Ages_10_17+Ages_18_44+Ages_45_64+RSP002_pneumonia+RSP010_aspiration_pneumonitis+RSP012_respiratory_failure+Use_mech_ventilation+ RateRespVisits+POPULATION, data = train)
summary(linreg2)

```

### In this second model we can see than most of variables pass the Hypothesis Test

```{r}
layout(matrix(c(1,2,3,4), 2, 2, byrow = TRUE))
plot(linreg2)
```
It seems than linear regressor model is not good to fit our dataset

### Now we calculate Root Mean Square Error (RMSE) to assess how well


```{r predict,message=FALSE,warning=FALSE}
Predic1<- predict(linreg2, newdata= test)
rmse(test$AQI,Predic1)
```

### RMSE value ≥0.5 reflects the poor ability of the model to accurately predict the data.

