---
title: "XGBoost"
author: "Marcos Mariscal"
date: "23/5/2021"
output: html_document
---
```{r setup, include=FALSE}

# Getting project directory 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
proydir <- rprojroot::find_rstudio_root_file()

```


#### This RMarkdown gaves last aproach with a XGBoost regression model

```{r call_libraries, include=FALSE}

# Calling R libraries
library(tidyverse)
library(ggplot2)
library(xgboost)
library(caret)
library(data.table)
library(DiagrammeR)

```


```{r envdefs, include=FALSE}
#Environment directories vars
airfilepath="./Data/Air_Quality/"
pngpath="./png/"
datapath="./Data/"

# Defining file names variables to load the data
file_TotRespData = "TotRespData.Rdata"

```



```{r Load_TotAgg, echo=FALSE}
# Read AQI Visits
load(file = paste(datapath,file_TotRespData, sep=""))


```

Getting only numeric Features for Respiratory illnesses

```{r numeric}

# taking out features not related to respiratory illnesses
NumDataResp <- TotRespData[,8:24]

```

---

#### Normalization the Data

---

Z-score normalization, consists of subtracting the mean and divide by the standard deviation
We normalize the data to bring all the variables to the same range

```{r norm}
means <- colMeans(NumDataResp)
sds <- apply(NumDataResp, 2, sd)

z_NumDataResp <- scale(NumDataResp, center = means, scale = sds)

z_NumDataResp <-  base::data.frame(z_NumDataResp)



```
### Column AQI is excluded because it will be our label column, the one we want to predict.

```{r}
df = z_NumDataResp[,!(names(z_NumDataResp) %in% "AQI")]
AQI_vector = z_NumDataResp$AQI
head(df)

```

### Splitting data into test and train

The following code splits 70% of the data selected randomly into training set and the remaining 30% sample into test data set.
train: will be used to build the model
test: will be used to assess the quality of our model.

```{r splitting}
set.seed(100)
ind <- sample(2, nrow(df), replace = T, prob = c(.7, .3))

df_train <- df[ind==1,1:16]
df_test <- df[ind==2, 1:16]

```

```{r labels}
t_train <- setDT(df_train)
t_test <- setDT(df_test)

labels <- df[ind==1, 16]
ts_labels  <- df[ind==2, 16]

```


```{r xgb}

dtrain <- xgb.DMatrix(label = labels, data = as.matrix(df_train))
dtest <- xgb.DMatrix(label = ts_labels, data = as.matrix(df_test))

```


### Extreme Gradient Boosting (xgboost). 
With linear model solver and tree learning algorithms.

eta=0.1; used in update to prevents overfitting (defult=0.3)
max_depth=15; the trees won’t be deep
nthread = 2; the number of CPU threads we are going to use
eval_metric="rmse"; Evaluation metrics for validation data, rmse for regression
nrounds=150;  the number of passes on the data. After these number of passes there is no rmse reduce

```{r modelxgboost, include=FALSE}
xgbmod <- xgboost(data = dtrain,
                  booster = "gblinear", 
               label = labels,
               eta=0.1,
               max_depth=15,
               nthread = 2,
               eval_metric ="rmse",
               nrounds = 500)

```

### Predicting

```{r predict}
# predict
print(xgbmod)

predic <- predict(xgbmod, dtest)

print(length(predic))

RMSE(predic, ts_labels)

```
#### Great RMSE value, less than 0.03


### Now visualize original test and predicted data in a plot

```{r visual}

x = 1:length(ts_labels)
plot(x, ts_labels, col = "red", type = "l")
lines(x, predic, col = "blue", type = "l")
legend(x = 50, y=6,  legend = c("original values", "predicted values"), 
       col = c("red", "blue"), box.lty = 2, cex = 0.8, lty = c(1, 1))
```

### Now we can see the Features importance

```{r Importance}
# Feature Importance
importance_matrix <- xgb.importance(model = xgbmod)
print(importance_matrix)
```

### Plot Importance Matrix

```{r AQI_Importance}
xgb.plot.importance(importance_matrix = importance_matrix)

```

```{r save}
#save model

saveRDS(xgbmod, file = "xgbmod.rds")
xgbmod<-readRDS("xgbmod.rds")
```
